---
title: "ds_18_coursework"
output:
  pdf_document: default
  html_document: default
date: "2023-11-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will work on a dataset containing a large subset of the human proteome, for each protein you are given its sequence, length, mass, and subcellular location (when available).

First, we import the file:
```{r}
library(readr)
df <- readr::read_csv("~/Downloads/uniref_merged.csv")
df
```

Next, we compute the number of each aminoacids from the sequence:
```{r}
library(stringr)
aminoacids <- unique(strsplit(paste(df$Sequence, sep="", collapse=""), '')[[1]])
# What aminoacids did we find? (You may note something quite interesting here...)
print(aminoacids)
for (c in aminoacids) {
  df[c] <- str_count(df$Sequence, c)
}
# Did we miss any amino acids?
all(apply(df[aminoacids], 1, sum) == df$Length)
```


Let's plot that!
```{r}
library(tidyr)
library(ggplot2)
pivot_longer(df[aminoacids], cols=aminoacids) %>%
ggplot(aes(x=name, y=value, fill=name)) +
  geom_boxplot()
```

Let's try a simple prediction, the mass as a function of the number of alanines:


```{r}
dfa <- df[c(aminoacids,"Mass")]
df.lm <- lm(Mass ~ A, dfa)
summary(df.lm)
```

We can now print the model, and check the coefficients:
```{r}
plot(dfa$Mass, predict(df.lm))
abline(a=1,b=1, col="red")
```
Can you make a near-perfect predictor for the mass, given the aminoacids count? (OK: R-squared > 0.9, Full credit: R-squared > 0.99):
```{r}

```

Split your data in a training and test set, and compute a summary statistic for the predictor quality:
```{r}

```


Our dataset also contains the subcellular location of proteins, stored as true/false values in the following columns: Golgi, Membrane, Extracellular, Endoplasmic Reticulum, Nucleus, Cytoplasm.

Can you create a plot showing the relationship between sequence length and subcellular location?
```{r}

```

Can you create a plot relating bias in sequence composition to subcellular location?
```{r}

```

Next, try to make a predictor for the subcellular location, given the sequence. (hint: start with the aminoacid counts!).
For full credit, your method should be able to predict multiple possible locations (with a precision greater than 0.2).
```{r}

```

Show the results of your predictions using a plot, and appropriate statistics. Are all the categories predicted with the same accuracy?
For full credit, the evaluation should be done on a test set separate from your training set.
```{r}

```

Finally, export your results.
For each predictor, export:
- a file with the real values, predicted values, and identifying information for each sample (each column should be appropriately labelled)
- a file with summary statistics, across predicted categories.
- a plot showcasing the prediction accuracy.
```{r}

```


For full credit, create an R shiny app, loading the exported files from the last step, and creating plots from it.
Do not attempt this unless you have successfully completed all the previous step!